Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-11-16 Revised-paper/Reply/Reply-Reviewer1.tex

Changes made in response to comments of Reviewer 1: \begin{enumerate}
%%1
\item
\textsf
{Abstract, hightlights, and introduction are written in a way that confuses the reader: the text jumps back and forth between bubbles and drops as well as shear flows and other flow configurations. Then it suddenly moves to lift and deformation rather than breakup. Finally a number of random applications are mentioned without any relation to bubble breakup. The authors should clearly state what their contrbution is from the outset. The reader will not go on until the end to figure that out by himself. The use of shorter paragraphs each addressing a single issue is very helpful.}
\vspace{3 mm}

Response: \\
In the revision, we made many improvements to the grammar and structure so that the reader can follow in a much more efficient manner. \\
\\

%%2
\item
\textsf
{Method: The use of only 2 levels of grid refinement appears odd. For an accurate representation of the flow near the interface a very fine mesh is typically needed. Only 32 cells over the bubble diameter is hardly sufficient ot get accurate results as corroborated by the mediorce agreement of the presented simulation results with the experimental data used for validation. Two more levels of grid refinement are likely needed to resolved the flow near the bubble while farther away some coarsening may be possible. Optimizing the grid is imperative in a situaltion like the considerd one due to the strong deformation of the bubble which high curvature of the interface and the singular nature of the breakup event. Fully capturing the breakup, which represents a topological singularity, is impossible without a subgrid scale model.}
\vspace{3 mm}

Response: \\
In the section entitled ``Selecting the appropriate grid size,'' we 
report results of a grid refinement study for our simulation results on
the threshold Reynolds number required for bubble break-up.  Based on
our grid refinement study, we claim that our results are within three
percent of the ``exact threshold'' assuming that we had access to an
infinite amount of computing power.  So, now the reviewer is asking for
more resolution in which we assert the reviewer request is 
unreasonable given the present state of technology.

As mentioned in our initial submission, and further emphasized in the revision,
if we refine the grid one more time, the simulation time will be 
lengthened by at least a factor 16! Why? (i) factor of 4 refinement 
due to 4 times as many grid points in the ``interface zone''; (ii) factor of $2^{3/2}$
refinement due to the surface tension stability time step constraint; (iii) factor of 2 refinement {\em OR MORE} because as one converges to the critical Reynolds number, the bubble oscillates, thus prolonging the ability to determine break-up versus no break-up.  Furthermore the bubble is longer at this critical juncture, thereby necessitating a longer and higher aspect ratio computational domain.  Further important notes: (a) we cannot just do adaptivity at the regions of high curvature because the resulting speed-up by using less grid points is offset by the further ill-conditioning of the matrix system in not having the interface wholly contained on the finest level.  As well, a sufficiently fine mesh is all around the gas-liquid interface in order to properly simulate the force exerted from the wall to the bubble. (b) The referee in his/her critism of our work did not mention alternate published work; and for good reason! Our work is the first of its kind! In the revised introduction, we make it clear that previous work on bubble deformation (just two papers) are based on the Lattice Boltzmann method which is an unacceptable method for measuring the tensile strength properties of a bubble.   (c) in our revised introduction, we make it clear why the bubble deformation problem is so much more difficult than the drop deformation problem and why our work is the FIRST to report on
tensile stress thresholding.  We have added further references to the introduction which understate the importance of our research on the ``tensile strength'' properties of a bubble, and at the same time we have made it clear why both controlled experiments and numerical simulations are so difficult.  In the Capillary and Reynolds number regime that we are interested in, experiments (which isolate the effect of density ratio) must be performed in microgravity, on the otherhand, computations are so much more difficult for the bubble case than the drop case because the density ratio is about 
1000 to one for the bubble case.

@@

%%3
\par\noindent
\item
\textsf
{Results: Mostly just pictures are shown from which little is to be learned. The only results of which future use can be expected is the regime map in Fig. 12. Bu this is also only qualitative and based on very few data points. More combinations of Ca and Re should be provided and a quantitative fit to the regime boundary be obtained. Moreover only two values of density and viscosity ratio equal to 1 and ~0.001 are used. An investigation of intermediate values would be straight forward and would allow to provide also models for the influence of these obviously important parameters. In particular it should be clarified whether the difference in behavior is more due to the density ration or more due to the viscosity ratio. Also it needs to be investigated how small these ratios need to be for the gas phase properties to be considered as negligibly small. For the practically important case of air bubbles in water the viscosity ratio is 0.02, an order of magnitude larger than considered in the paper. Whether this can already be considered negligible or not remains unclear.}
\vspace{3 mm}

Response: \\
As in our response to the previous comment, we believe it is unreasonable to report more results; specifically each new data point requires an additional year, and additional computers.  As also mentioned in our response to point 2 above, and hilighted in our revised introduction,
our work as it stands now is significant and novel.  We take the referee comments on as a challenge for future work.

We remark that the significance of our article is the fact that we have 
reported for the first time controlled tensile strength properties for a bubble in shear flow.  In addition, we have made the following noteworthy discoveries along the way:
\begin{enumerate}
\item The behavior of the bubble deformation process around critical $Re$ conditions with $Ca$ = 0.3  $\sim$ 1.0 has been presented for the first time.
\item The behavior of the bubble breakup process at critical $Re$ conditions with $Ca$ = 0.3  $\sim$ 1.0 has been presented for the first time.
\item The value of critical $Re$ numbers was found for $Ca$ = 0.3  $\sim$ 1.0 for the first time.
\item  We showed that most previous computational studies overestimated drop deformation and breakup because most previous studies used a domain size with $W$ (width ($y$directional length)) =4$R$. Drop deformation is promoted when using a domain size with $W=4R$. 
\item The difference in the deformation and breakup between the bubble ($\lambda<<1$) and the drop ($\lambda = 1$ and $\eta = 1$) was clearly shown.
\item Shear-induced bubble/drop breakup critical curves were obtained. 
\end{enumerate}

Regarding the effect of $\lambda$ and $\eta$, ongoingly we are examining the effect of $\lambda$ and $\eta$ on drop breakup for following 
combinations of $\lambda$ and $\eta$:

### 2024-11-16 Revised-paper/Reply/Reply-Reviewer2.tex

%%2
\item
\textsf
{Page 2 Line 14: There should be an "and" before "Renardy"}
\vspace{3 mm}

Response: \\
We appreciate the reviewer's suggestion. The description of the applicable part disappeared 
in the revised paper because we have completely modified the ``Introduction".\\
\\
Please see the ``Introduction".
\\

%%3
\par\noindent
\item
\textsf
{The authors mentioned several DNS methods in the Introduction section while choosing the CLSVOF method in this work. What are the reasons?}
\vspace{3 mm}

Response: \\
The level set method, by itself, is not designed to directly conserve the
volume of materials.  The error is especially large at regions of high curvature.  The lack of volume conservation can significantly alter the numerical predictions.  On the other hand, the volume of fluid method, by itself, is not designed to accurately and robustly approximate the interface curvature.  The surface tension force will be incorrectly calculated if the curvature is not accurately approximated.  We have found that the CLSVOF method on the otherhand, in which we couple the two methods, is the optimal method.  We have added further discussion to the paper.

\\
\\

@@

%%4
\par\noindent
\item
\textsf
{Why can the effect of gravity be disregarded in this work?\\}
\vspace{3 mm}

Response: \\
In our computations, we set $g = 0$ because we wanted to isolate only the effects of the density and viscosity ratios and 
wanted to compare with previous studies dealing with the deformation and breakup of a drop with $\lambda =1$ and $\eta = 1$. 
Also, in the revised paper, a brief discussion of the potential influence of gravity (Froude number) on bubble breakup dynamics 
has been added according to another reviewer's suggestion.\\
\\
Please see lines 133 to 139.
\\

### 2024-11-16 Revised-paper/Paper/Bubble-deformation-Rev.tex
need to proofread the paper

between gas and liquid is tracked with a coupled level set/volume-of-fluid (CLSVOF) interface capturing algorithm \cite{SusPuc00,SusSmiHusOhtZhi07}.  

@@

flow\cite{WeiQiaXu12,WanShiZha15,AAMM-16-2}  use the (explicit) Lattice Boltzmann method.  For accurately computing the tensile strength of a bubble, and accurately computing threshold parameters for break-up (what we do in this article, and what was not done in previous work), it is critical that a numerical method directly enforces the velocity continuity condition and the gas-liquid interface normal jump conditions.  We contend that a projection method (e.g. this paper and \cite{zhang2021three,zhang2022three,OhtSus12}) is the more appropriate (albeit slower) method for our study rather than the Lattice Boltzmann method.  Also, in contrast to the Lattice Boltzmann method, our interface ``capturing'' method, the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07},  maintains the gas/liquid interface as sharp and by construction our method preserves mass and volume within a fraction of a percent. The results that we present in this article (see e.g. section \ref{convergence}) regarding measuring the tensile strength of a bubble are unique and validated with respect to comparisons with previous experimental data and grid refinement studies.  Admittadly, each simulation on the finest resolution takes over a half a year to complete on a workstation because of the following unavoidable factors: (i) the large density-ratio projection method requires the solution of a large sparse, ill-conditioned, matrix system at each time step, (ii) the finer the mesh, the more precise the measured threshold, and right at the threshold (Taylor Deformation parameter $D\approx 1$!), oscillatory behaviour is observed delaying the determination of breakup or not.  Finally, the larger the deformation parameter $D$, the longer we must make our computational domain (bigger domain aspect ratio!) thereby adversely effecting the condition number even more for carrying our the pressure projection. \\ \indent

@@

When comparing with previous drop studies, we fix $\lambda=1$.  As a result,
(for $\lambda=1$) the following dimensionless physical parameters are used to
describe the problem of drop deformation/breakup in shear flow
%
\begin{eqnarray}\label{dimensonless}
  Re = \frac{\rho_{\rm m}UR}{\mu_{\rm m}}, \quad
  Ca = \frac{\mu_{\rm m}U}{\sigma}, \quad
  \eta = \frac{\mu_{\rm b}}{\mu_{\rm m}} 
	\hspace{20pt} m=\mbox{matrix fluid}
\end{eqnarray}
%
$U$ is the velocity scale and $\sigma$ denotes the surface tension.  For the
problem of shear-induced drop deformation and breakup, the velocity is set to
$U = \mathit{\Gamma} R$, where the shear-rate is $\mathit{\Gamma} = 2V/H$.  As
mentioned in the introduction, most previous drop studies set $\eta = 1$ (e.g.
\citet{LiRenRen00}).  Thus, for comparison with previous drop deformation and
breakup problems, we set $\lambda = \eta = 1$ (and also neglect the effect of
gravity so that $g=0$). On the other hand, in our computations for bubble
deformation, we set the density and viscosity of air to be $\rho_{\rm b}$ = 1.2
kg/m$^{3}$ and $\mu_{\rm b} = 1.8 \times 10^{-5}$ Pa$\cdot$s respectively.  We
emphasize that for consistency with previous studies (\citet{LiRenRen00,
RusMan02, MulTobDreFisWin08, KomShaEskDer14, AmaBalCasOli19}), we
computationally examine the deformation and breakup of a bubble in simple
linear shear flow as a function of the $Re$ and $Ca$ numbers.  That is to say,
by setting $g=0$, we are isolating the effect of only varying $Re$ and $Ca$ on
bubble deformation and breakup.  In our controlled study, we determine the
critical $Re_{\rm c}$ versus $Ca$ curve in which $Re_{\rm c}$ corresponds to
the threshold of bubble (or drop) breakup.  We determine the critical $Re_{\rm
c}$ versus $Ca$ curve for strategic pairs of the density ratio and viscosity
ratio.

@@

The temporal discretization of our numerical method is an operator split
projection method as described by \citet{SusSmiHusOhtZhi07}.  An outline of 
our method is as follows (see \citet{SusSmiHusOhtZhi07}, section 4, for more 
details):
\begin{description}
\item[Timestep $\Delta t$] The timestep is governed by the CFL condition and
  surface tension (section 5.7 of \citet{SusSmiHusOhtZhi07}):
 \begin{eqnarray*}
 \Delta t < \min_{i,j,k} \left( \frac{\Delta x}{2|U^{n}|},
   \frac{1}{2}\sqrt{\frac{\rho^{L}}{8\pi\sigma}}\Delta x^{3/2}\right)
 \end{eqnarray*}

@@

To make efficient use of computational resources, our numerical simulations
utilize an adaptive hierarchy of grids based on a dynamic adaptive mesh
refinement (AMR) technique (\citet{SusAlmBelColHowWel99}).  Adaptive grids are
dynamically adjusted based on the location of the 
deforming gas-liquid interface.
In the AMR technique the grid resolution is increased in regions near the
interface while a coarser grid is used where the flow is relatively steady.
The upper panel of Figure~\ref{fig:SchemAndGrid}(b) displays a schematic view
of the hierarchical grid structure and the lower panel corresponds to an actual
computational example for bubble deformation in simple linear shear flow.  In
general, the mesh hierarchy is composed of different levels of refinement
ranging from coarsest $\ell=0$ (``level-0") to finest
$\ell=\ell_{\textrm{max}}$ (``level-$\ell_{\textrm{max}}$").  The refinement
ratio of one grid size ($\Delta x=\Delta y=\Delta z$) to the next finer level
is two so that $\Delta x^{\ell+1}=0.5\Delta x^{\ell}$.  All computations in
this study used an AMR system with a maximum prescribed level
$\ell_{\textrm{max}} = 1$ (as illustrated in the upper panel of
Figure~\ref{fig:SchemAndGrid}(b)).  In our adaptive mesh refinement algorithm,
the velocity in coarse grid cells that neighbor fine grid cells is interpolated
from the coarse grid using bilinear interpolation in order to initialize
``ghost'' fine cells. Thus, the bilinear interpolation procedure produces
interpolated fine grid data as a linear combination of the coarse grid data.

>>>>

between gas and liquid is tracked with a coupled level set/volume-of-fluid (CLSVOF) interface capturing algorithm \cite{SusPuc00,SusSmiHusOhtZhi07}.  The rationale for the CLSVOF method is that the hybrid method represents the (complex) gas-liquid interfaces with minimal volume loss (property of the volume-of-fluid method) and minimal error in the approximation of the surface tension force (property of the level set method).

@@

flow\cite{WeiQiaXu12,WanShiZha15,AAMM-16-2}  use the (explicit) Lattice Boltzmann method.  For accurately computing the tensile strength of a bubble, and accurately computing threshold parameters for break-up (what we do in this article, and what was not done in previous work), it is critical that a numerical method directly enforces the velocity continuity condition and the gas-liquid interface normal jump conditions.  We contend that a projection method (e.g. this paper and \cite{zhang2021three,zhang2022three,OhtSus12}) is the more appropriate (albeit slower) method for our study rather than the Lattice Boltzmann method.  Also, in contrast to the Lattice Boltzmann method, our interface ``capturing'' method, the CLSVOF method\cite{SusPuc00,SusSmiHusOhtZhi07},  maintains the gas/liquid interface as sharp, enables accurate approximation of the surface tension force,  and by construction the CLSVOF method preserves mass and volume within a fraction of a percent. The results that we present in this article (see e.g. section \ref{convergence}) regarding measuring the tensile strength of a bubble are unique and validated with respect to comparisons with previous experimental data and grid refinement studies.  Admittadly, each simulation on the finest resolution takes over a half a year to complete on a workstation because of the following unavoidable factors: (i) the large density-ratio projection method requires the solution of a large sparse, ill-conditioned, matrix system at each time step, (ii) the finer the mesh, the more precise the measured threshold, and right at the threshold (Taylor Deformation parameter $D\approx 1$!), oscillatory behaviour is observed delaying the determination of breakup or not.  Finally, the larger the deformation parameter $D$, the longer we must make our computational domain (bigger domain aspect ratio!) thereby adversely effecting the condition number even more for carrying our the pressure projection. \\ \indent

@@

When comparing with previous drop studies, we fix $\lambda=1$ for the drop cases.  As a result,
(for $\lambda=1$) the following dimensionless physical parameters are used to
describe the problem of drop deformation/breakup in shear flow
%
\begin{eqnarray}\label{dimensonless}
  Re = \frac{\rho_{\rm m}UR}{\mu_{\rm m}}, \quad
  Ca = \frac{\mu_{\rm m}U}{\sigma}, \quad
  \eta = \frac{\mu_{\rm b}}{\mu_{\rm m}} 
	\hspace{20pt} m=\mbox{matrix fluid}
\end{eqnarray}
%
$U$ is the velocity scale and $\sigma$ denotes the surface tension.  For the
problem of shear-induced drop deformation and breakup, the velocity is set to
$U = \mathit{\Gamma} R$, where the shear-rate is $\mathit{\Gamma} = 2V/H$.  As
mentioned in the introduction, most previous drop studies set $\eta = 1$ (e.g.
\citet{LiRenRen00}).  Thus, for comparison with previous drop deformation and
breakup problems, we set $\lambda = \eta = 1$ (and also neglect the effect of
gravity so that $g=0$). On the other hand, in our computations for bubble
deformation, we set the density and viscosity of air to be $\rho_{\rm b}$ = 1.2
kg/m$^{3}$ and $\mu_{\rm b} = 1.8 \times 10^{-5}$ Pa$\cdot$s respectively.  We
emphasize that for consistency with previous studies (\citet{LiRenRen00,
RusMan02, MulTobDreFisWin08, KomShaEskDer14, AmaBalCasOli19}), we
computationally examine the deformation and breakup of a bubble in simple
linear shear flow as a function of the $Re$ and $Ca$ numbers.  That is to say,
by setting $g=0$, we are isolating the effect of only varying $Re$ and $Ca$ on
bubble deformation and breakup.  In our controlled study, we determine the
critical $Re_{\rm c}$ versus $Ca$ curve in which $Re_{\rm c}$ corresponds to
the threshold of bubble (or drop) breakup.  We determine the critical $Re_{\rm
c}$ versus $Ca$ curve for strategic pairs of the density ratio and viscosity
ratio.

@@

The temporal discretization of our numerical method is an operator split
projection method as described by \citet{SusSmiHusOhtZhi07}.  An outline of 
our method is as follows (see \citet{SusSmiHusOhtZhi07}, section 4, for more 
details):
\begin{description}
\item[All Steps. Timestep] 
 \begin{eqnarray*}
 \end{eqnarray*}
  The timestep, $\Delta t$, is governed by the CFL condition and
  surface tension (section 5.7 of \citet{SusSmiHusOhtZhi07}):
 \begin{eqnarray*}
 \Delta t < \min_{i,j,k} \left( \frac{\Delta x}{2|U^{n}|},
   \frac{1}{2}\sqrt{\frac{\rho^{L}}{8\pi\sigma}}\Delta x^{3/2}\right)
 \end{eqnarray*}

@@

To make efficient use of computational resources, our numerical simulations
utilize an adaptive hierarchy of grids based on a dynamic adaptive mesh
refinement (AMR) technique (\citet{SusAlmBelColHowWel99}).  Adaptive grids are
dynamically adjusted based on the location of the 
deforming gas-liquid interface.
In the AMR technique the grid resolution is increased in regions near the
interface while a coarser grid is used where the flow is relatively steady.
The upper panel of Figure~\ref{fig:SchemAndGrid}(b) displays a schematic view
of the hierarchical grid structure and the lower panel corresponds to an actual
computational example corresponding to bubble deformation in simple linear shear flow.  In
general, the mesh hierarchy is composed of different levels of refinement
ranging from coarsest $\ell=0$ (``level-0") to finest
$\ell=\ell_{\textrm{max}}$ (``level-$\ell_{\textrm{max}}$").  The refinement
ratio of one grid size ($\Delta x=\Delta y=\Delta z$) to the next finer level
is two so that $\Delta x^{\ell+1}=0.5\Delta x^{\ell}$.  All computations in
this study used an AMR system with a maximum prescribed level
$\ell_{\textrm{max}} = 1$ (as illustrated in the upper panel of
Figure~\ref{fig:SchemAndGrid}(b)).  In our adaptive mesh refinement algorithm,
the velocity in coarse grid cells that neighbor fine grid cells is interpolated
from the coarse grid using bilinear interpolation in order to initialize
``ghost'' fine cells. Thus, the bilinear interpolation procedure produces
interpolated fine grid data as a linear combination of the coarse grid data.
